On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-02-24 MS.tex
LIzzy comments 1

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 307/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate is seen as a central element of its ecological strategy, and as determined by its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and growth, casting doubt on the capacity of some prominent traits to influence growth rate.

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theoretical work, and test these predictions in tree species using a meta-analysis of 103 studies ($>$ 500 correlations) for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate). We also recorded data for 14 other traits commonly used in the trait literature. To capture the effects of plant size, we tested for a shift in the direction of correlation between growth rates and traits across three ontogenetic stages: seedling, sapling and adult.

@@

\item Synthesis: We show how trait-growth correlations change in a predictable way with plant size. Our understanding of plant strategies should shift away from attributing species as having slow or fast growth throughout their life, in favour of attributing growth trajectories.
\end{enumerate}

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients, and light for growth; and then ultimately replacing themselves with offspring. The idea that strategy differences among species might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who observed a wide spread of relative growth rate  among seedlings growing under favourable conditions. Seedling potential relative growth rate has continued to be used as a major strategy indicator within Grime's CSR (Competitor -- Stress tolerator -- Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling relative growth rate into different components showed that specific leaf area (SLA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. During the 1990's a new style of ``trait ecology'' arose whereby measurable traits, including SLA, were used directly as strategy axes \citep{Westoby:2002ft}. This approach, using measurable traits, made broad comparisons possible across continents, latitudes, and thousands of species. Since then, functional traits have been seen as constructive tools to understanding plant growth strategies \citep{Westoby:2002ft}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

In this paper, we consider five prominent traits (specific leaf area = SLA, wood (or stem) tissue density, maximum leaf photosynthetic capacity per unit area = Aarea, maximum (or asymptotic height), and seed mass) and we investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
Using a meta-analysis approach, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. We anticipated a sizeable literature on traits and growth to be available but we were uncertain about the representation of all trait-stage combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between functional traits and plant growth.

@@

Assuming plants otherwise maintain a similar cross section of conducting tissue, cheaper stem construction (low wood density or stem density) will increase the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$), and consequently growth rate in height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all ontogenetic stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between wood density and growth may also strengthen in larger plants. A similar shift is expected for growth in mass growth, in either absolute or relative terms. Note in the case of mass growth, the effect of cheaper stem construction is realised not in the $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ term , but via greater total leaf area at a given mass, i.e. the accumulated effect of past reductions in  $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$.

@@

The only effect of increased seed size in the above model is to produce a larger overall seedling. Under similar light conditions, larger seedlings are therefore predicted to have faster absolute growth rates . However, because the ratio of leaf area to total biomass decreases with plant size, larger seedlings are predicted to have lower mass-based relative growth rates during the earliest phases of growth \citep{Turnbull:2012ew}(Table \ref{tab:trade-offs}). Yet, as plants grow, we expect the effect of this initial size difference to be minimal, and the correlation between seed size and growth rate should therefore disappear. For larger plants, seed mass is not expected to have any distinct influence on growth rates (Table \ref{tab:trade-offs}).

@@

We used categorical groupings by ontogenetic ``stage'' to test our hypotheses regarding the effects of traits on growth rates. The authors of the 103 articles used in this analysis reported their plants as belonging to any of five different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. Because these categories were established according to the expertise of the authors, we were concerned about the risk of assigning plants of similar size to different categories when compiling data across multiple studies and vegetation types. This risk turned out to be small: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages were relatively compatible (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m). There was also a clear break between saplings and adults in diameter (at 10 cm). Comparing these values to the bivariate relationships observed between variables (age, height, diameter) in the ``Biomass And Allometry Database'' \citep{Falster:2015}, demonstrated that the thresholds identified above for each stage transition also aligned relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries.

@@

We thank Andrea Stephens for helpful comments on the meta-analyses, and Sean Gleason et al for pre-publication access to their data. Charles Warren and Jordi Mart{\'{i}}nez-Vilalta kindly provided additional information about their published results. This work was funded by the Australian Research Council through a fellowship to M.W, a discovery grant to D.S.F and a Macquarie University Research Excellence Scholarship and L'Oreal UNESCO FWIS Fellowship to E.F.G.

>>>>

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 309/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate is seen as a central element of its ecological strategy, and as determined by its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and growth, casting doubt on the capacity of some prominent traits to influence growth rate.

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theoretical work, and test these predictions in tree species using a meta-analysis of 103 studies ($>$ 500 correlations) for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate). We also recorded data for 14 other traits commonly used in the trait literature. To capture the effects of plant size, we tested for a shift in the direction of correlation between growth rates and each trait across three ontogenetic stages: seedling, sapling and adult.

@@

\item Synthesis: We show for some traits, how trait-growth correlations change in a predictable way with plant size. Our understanding of plant strategies should shift away from categorizing species as having slow or fast growth throughout their life, in favour of attributing growth trajectories.
\end{enumerate}

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients, and light for growth; and then ultimately replacing themselves with offspring. The idea that strategy differences among species might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who observed a wide spread of relative growth rates among seedlings growing under favourable conditions. Seedling potential relative growth rate has continued to be used as a major strategy indicator within Grime's CSR (Competitor -- Stress tolerator -- Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling relative growth rate into different components showed that specific leaf area (SLA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. During the 1990's a new style of ``trait ecology'' arose whereby measurable traits, including SLA, were used directly as strategy axes \citep{Westoby:2002ft}. This approach, using measurable traits, made broad comparisons possible across continents, latitudes, and thousands of species. Since then, functional traits have been seen as constructive tools to understanding plant growth strategies \citep{Westoby:2002ft}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

In this paper, we consider five prominent traits (specific leaf area = SLA, wood or stem tissue density, maximum leaf photosynthetic capacity per unit area = Aarea, maximum or asymptotic height, and seed mass) and we investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
Using a meta-analysis approach, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. We anticipated a sizeable literature on traits and growth to be available but we were uncertain about the representation of all trait-stage combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between functional traits and plant growth.

@@

Assuming plants maintain a similar cross section of conducting tissue, cheaper stem construction (low wood density or stem density) will increase the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$), and consequently growth rate in height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all ontogenetic stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between wood density and growth may also strengthen in larger plants. A similar shift is expected for growth in mass, in either absolute or relative terms. Note in the case of mass growth, the effect of cheaper stem construction is realised not in the $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ term , but via greater total leaf area at a given mass, i.e. the accumulated effect of past reductions in  $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$.

@@

The only effect of increased seed size in the above model is to produce a larger overall seedling. Under similar light conditions, larger seedlings are therefore predicted to have faster absolute growth rates. However, because the ratio of leaf area to total biomass decreases with plant size, larger seedlings are predicted to have lower mass-based relative growth rates during the earliest phases of growth \citep{Turnbull:2012ew}(Table \ref{tab:trade-offs}). Yet, as plants grow, we expect the effect of this initial size difference to be minimal, and the correlation between seed size and growth rate should therefore disappear. For larger plants, seed mass is not expected to have any distinct influence on growth rates (Table \ref{tab:trade-offs}).

@@

We used categorical groupings by ontogenetic ``stage'' to test our hypotheses regarding the effects of traits on growth rates. The authors of the 103 articles used in this analysis reported their plants as belonging to any of five different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. Because these categories were established according to the expertise of the authors, we were concerned about the risk of assigning plants of similar size to different categories when compiling data across multiple studies and vegetation types. This risk turned out to be small: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages were relatively compatible with thresholds we defined (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m). There was also a clear break between saplings and adults in diameter (at 10 cm). Comparing these values to the bivariate relationships observed between variables (age, height, diameter) in the ``Biomass And Allometry Database'' \citep{Falster:2015}, demonstrated that the thresholds identified above for each stage transition also aligned relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries.

@@

We thank Andrea Stephens for helpful comments on the meta-analyses, and Sean Gleason et al for pre-publication access to their data. Charles Warren and Jordi Mart{\'{i}}nez-Vilalta kindly provided additional information about their published results. Elizabeth Wenk and Freya Thomas for their friendly review of our manuscript. This work was funded by the Australian Research Council through a fellowship to M.W, a discovery grant to D.S.F and a Macquarie University Research Excellence Scholarship and L'Oreal UNESCO FWIS Fellowship to E.F.G.
